I would like to express my sincere gratitude to my supervisor, Associate Professor Akira Furui, for his constant and enthusiastic guidance and great encouragement throughout the course of this research.
Associate Professor Furui provided me with detailed and careful instruction on every aspect of my work, ranging from discussions on the direction of the research to thesis writing and presentation techniques.
Although I had mixed feelings of anticipation and anxiety when I was first assigned to the laboratory, thanks to his advice—at times kind and at times strict—I was able to push forward with my research without hesitation.

I would like to express my deep appreciation to Professor Hiroaki Mukaidani and Associate Professor Zu Soh, who served as my sub-supervisors,
for their valuable advice during the interim presentation. The insightful comments and sharp feedback I received from them served as extremely important guidelines for improving the quality of this research and deepening the discussion.

I am also deeply grateful to Assistant Professor Hiroaki Aizawa for providing valuable opinions from multiple perspectives during our collaborative research.
He taught me how the fundamental knowledge gained in undergraduate lectures connects to cutting-edge research issues, giving me a rare and invaluable opportunity to reaffirm the depth and fascination of academia.

The presence of the members of the Intelligent Biological Information Systems Laboratory was a great emotional support for me in my daily research life.
The time spent not only in active discussions during the general seminars but also sharing meals between research activities and laughing over casual daily conversations are irreplaceable memories for me.

Finally, I would like to dedicate my heartfelt thanks to my parents, who willingly approved my pursuit of graduate studies and continued to support me both financially and mentally throughout my enrollment.
The privileged environment that allowed me to devote myself to research without any inconvenience was realized solely because of the devoted support of my family.

I hereby express my deepest gratitude.